Analyzing the impact of social attributes on the success of commit integration

### 2017-02-18 Paper/MSRChallenge2017.tex
after Mau's pass

\begin{abstract}
One important aspect of developing open source software is deciding which
commits to include in the main branch of a project.  
Recent research has started to explore how the open source community manages
this question and found that both technical and social factors are factors in 
the success of project contributions on GitHub.  
This investigation defined a successful project contribution as a project
contribution that gets
accepted to the project.  However, with the continuous integration, we can
explore this question a step further.  We can better estimate the success of a
contribution by investigating if the contribution passed the project's
test cases.  We performed an initial investigation into the technical and
social factors involved in a commit that successfully passes the integration
tests. Specifically, we investigated if social factors (such as being a
core member of the development team, having a large number of followers,
or contributing a large number of commits) improve predictions of build
success.  We have found that while social factors cause an noticeable increase
in predictive power (12\%), further investigation into these factors are
required.
\end{abstract}

@@

\section{Introduction}
For managers of an open source project, it is important to ensure that the
project consists of high quality contributions.  The decision to include code in
a project has been described as a meritocracy, where only the code is a factor
in the decision~\cite{Scacchi07}.  However, to better understand successful 
project contributions, we must analyze factors which are common to all commits.

Over the years, the open source
community has developed various techniques to vet project contributions, such as pull
requests which must be approved by a core team member~\cite{gousios14}. One such
technique is continuous integration. Continuous integration automatically checks
the contribution against the project's test cases, to notify the development
team if the change caused a test case failure~\cite{Meyer14}.  When a change
passes one of these vetting techniques, it shows that the project developers
have confidence in the changes.

In past studies, both technical factors and social factors have been found to be
important when when predicting pull request success~\cite{tsay14icse,tsay14fse}.
Yet, these factors may not be as important during continuous integration.
Specifically, we want to investigate if social factors, which can be used a
proxy for experience, are important when predicting build success when grouped
with technical data, used as a proxy for commit complexity.

@@

\textbf{Related work:}
Similar studies have shown the relationship between different attributes and 
different measures of success: pull request acceptance~\cite{gousios14, 
gousios15,tsay14icse,tsay14fse}, pull request evaluation latency~\cite{Yu15}.
Tsay et al~\cite{tsay14icse} performed a very relevant study where they 
determine that the social connection between the developer who is asking the 
pull request and the integrators is a key factor torwards acceptance of a pull 
request. Unlike this previous study, in our study we look at a broader aspect of 
social aspects and how this attributes towards the success in Travis CI jobs. 
Gousios et al~\cite{gousios14} study the relationship between the area of the 
source code being modified by the developer requesting the changes to be pulled 
and the acceptance rate of these changes is highly correlated. Also existence of 
tests and overall code quality~\cite{gousios15} are attributes that have a 
strong relationship with the acceptance rate of a pull request. Bettenburgh et 
al~\cite{bettenburgh10} discuss the impact of social interactions with 
post-release defects, Vasilescu et al~\cite{vasilescu14} explored how different 
project characteristics (programming language, project age, etc.) affect commit 
success when integrated to the automatic build process. Yu et al~\cite{yu16} 
analyze the impact of CI usage in Github and its correlation with software 
quality. Unlike the works previously listed, our study focuses on independent social 
attributes and how these relate to the success of the integration of a commit 
pushed by the developer being described by these social attributes.

Our study shows that social factors are an important aspect of build success but
these factors alone do not fully predict if a build will succeed.  We also show
that core team members are more likely to achieve build success than non-core
team members and having followers increases the likelihood passing the
continuous integration build.

@@

\subsection{Goal}
The goal of this study is to have a deeper understanding of the relationship 
between the several different social properties that that estimate the
developer's experience, and the likelihood of a developer's commit
to be considered a successfully integrated in the project. We define successful
integration as passing the continuous integration build tests in the project.

@@

Previous studies have identified a relationship between technical and social 
factors and pull request acceptance. We are interested in understanding if
social factors, that proxy experience, can be used to predict
a different success measure: continuous integration success.

@@

\subsection{Metric}
We created a linear regression model based on two different sets of data. Merely 
technical data related to technical aspects of the complexity of the commit, and 
a superset of this before-mentioned data and social data gathered from the 
developer who pushed the commit. We evaluate the predictive power of these two 
models with 10 fold cross validation and we analyze which of them is able to 
predict with higher accuracy the success of unseen commits.

\section{Building Dataset}
\label{approach}
To investigate our research questions, we built a dataset based on the 
attributes that consisted of both technical and social factors.  We collected
the technical factors from the provided data set~\cite{msr17challenge}.
We extracted only the attributes in the data set that proxy the commit's
complexity and the build status.
\begin{itemize}
%\item tr\_original\_commit
%\item gh\_build\_started\_at 
\item git\_diff\_src\_churn 
\item gh\_diff\_files\_added 
\item gh\_diff\_files\_deleted 
\item gh\_diff\_files\_modified 
\item gh\_diff\_tests\_added 
\item gh\_diff\_tests\_deleted 
\item gh\_diff\_src\_files 
\item gh\_sloc 
\item tr\_status 
\end{itemize}

Once we had these attributes, we removed repeated rows \todo{Why did we have to
remove repeated rows?}

To collect the social data, we created a web scrapper to extract the social data
from developer's profiles.
We used the commit hash and the project name in the original data
set~\cite{msr17challenge} to locate the web page of the commit using the syntax
below. 
\newline \newline
\textit{https://github.com/\textbf{NameOfProject}/commit/\textbf{CommitHash}}
\newline \newline
Then, we scraped the commit author's
\textit{username} from the page. If the scrapper was able to find username, the
scraper went to the user's profile information using the syntax below.
\newline \newline
\textit{https://github.com/\textbf{UserName}}
\newline \newline
The scrapper then extracted the following items from the user's profile.
\begin{itemize}
\item Repositories
\item Stars
\item Followers
\item Following
\item Contributions 
\end{itemize}

The dataset used in this study~\cite{msr17challenge} is very diverse 
and consists of commits that date from far back, and the social data we are 
gathering is recent data, gathered in January 2017. The researchers of this 
project acknowledge this discrepancy and in an effort to mitigate the threat of
using future result to predict previous events,
we removed the commits in the dataset that were not performed within the last
three years of the data set's release (2014,2015,2016).  While this is not
perfect, we believe that these numbers are unlikely to have changed
significantly in the last three years.

After building the data set, we converted the various Travis build statuses to
successful (passed in the data set) or not successful (errored, failed, canceled
and started in the data set).  This still left a big discrepancy in the two
categories, with about a 3:1 successful to not successful ratio. To avoid having
this discrepancy alter our results, we created a separate data set where the
successful builds were randomly selected until the number of successful builds
equaled the number of not successful builds (downsampled\todo{citation needed}).  We used both the
downsampled and non-downsampled data sets in our results.

We evaluated the predictive power of the social factors using a decision
tree~\cite{Quinlan86}.
Decision trees are a predictive model which uses binary decisions on the
attributes to predict a final class.  The predicted class is represented as the
leaves, while the attribute decisions used in the prediction cause the
different branches.
In this study, we used the Weka~\cite{Weka} J48~\cite{Quinlan1993} decision tree
with default settings. The decision trees were evaluated using ten fold cross 
validation with the data in each data set.

\subsection{Success Classification}
\begin{table*}[ht]
\centering
\begin{tabular}{l||r|r||r|r}
  & \multicolumn{2}{|c||}{Original number of instances}  & \multicolumn{2}{|c}{Downsampled} \\
\hline
  & Technical & Technical and Social & Technical & Technical and Social \\
\hline
\textbf{Correctly Classified Instances}&  77.01\% & 81.10\%  & 64.10\% & 76.09\%\\

@@

\hline
\end{tabular}
\\
\center
  \caption{ Second and third columns: Comparison of performance when running 10 fold cross validation of two decision trees. The first column after the first double bar uses only technical attributes to predict the Travis Job status, the second column augments the dataset with social attributes. Fourth and fifth columns: The same intuition as the previous two, with the difference that the dataset was downsampled to provide the same number of successful and unsuccessful instances }
  \label{resultsTable}
\end{table*}

@@

Table~\ref{resultsTable} shows the results of running the two data sets through the
default Weka J48 decision tree algorithm.The decision tree built from
only the technical indicators has a correct classification rate of 77.01\% which
means the technical attributes only increased the prediction accuracy by about 3\%.
Adding the social data increased the classifier's accuracy from 77.01\% to
81.10\%, about a 4\% increase.

@@

\section{Discussion}
We have found that social attributes which proxy experience produce a small but
noticeable increase in predicting build success.  While the increase was very
small with the full data set, there was a greater increase in prediction
accuracy after performing downsampling.  These results show that experience is
important to predicting build commit success but a strong predictor.

The build success by core team member results show that core team members are
more likely to have the build integration pass than non-core team members.  This
result provides evidence that non-core team member commit are not throughly
scrutinized, although further studied will be required to draw definite
conclusions.

@@

\section{Threats to validity}
This study contains various threats to validity.

\noindent\textbf{Internal validity:}\\
One source of internal validity concern is the time mismatch between the
technical and social data.  We restricted the commits to only the commits in the
last three years of the data set to reduce this concern.
There is also a chance that scripts we used to create the data sets contain an
error.  For full transparency, we've released the scripts used to collect the
data set at (This will be added after the anonymous review).

%footnote{https://goo.gl/CqTX44}

\noindent\textbf{External validity:} \\
While we have used a data set with high quality standards~\cite{msr17challenge},
these results may not generalized to non-open source projects.

\noindent\textbf{Construct validity:}\\
While we believe the build status is a good metric for contribution success,
there is the chance that a problem in the project's test cases, and not the
contribution itself, led to the build failure.  There are also other possible
technical and social factors which were not explored in this study.  Future
research will be required to determine if those factors are important.

\section{Conclusion}
In this study we have analyzed the ability to use social factors to predict
build integration success.
Our results show that adding social data, which proxies experience, provides a
small increases in prediction accuracy.  This leads us to conclude that while
experience is important, it is not the only factor in build success.
We have also found that non-core contributors cause not successful builds more
than core contributors.  This leads us to believe that non-core contributions
are not inspected as closely as they should be.  We have also found that having
a follower, and especially a large follower base, can be an indicator of build
integration success.

>>>>

\begin{abstract}
Software quality is an important and desirable aspect of the software 
development process overall, especially with the increasing participation of a very
diverse pool of developers, the number of commits and pull requests keeps increasing 
and the mechanisms available to identify which developers are pushing high 
quality commits and pull requests is scarce.
Recent research has explored how the open source community manages
this question finding that both technical and social factors are factors in 
the success of project contributions on GitHub.  
This investigation defined a successful project contribution as a project
contribution that gets
accepted to the project. However, with continuous integration, we can
explore this question a step further.  We can better estimate the success of a
contribution by investigating if the contribution passed all the quality standards set 
by the repository owners.  We performed an initial investigation into the technical and
social factors involved in a commit that successfully passes the integration
tests. Specifically, we investigated if social factors (such as being a
core member of the development team, having a large number of followers,
or contributing a large number of commits) improve predictions of build
success.  We have found that while social factors cause a noticeable increase
in predictive power (12\%), further investigation into these factors are
required.
\end{abstract}

@@

\section{Introduction}
For managers of an open source project, it is important to ensure that the
project consists of high quality contributions.  The decision to include code in
a project has been described as a meritocracy, where only the code is a factor
in the decision~\cite{Scacchi07}.  However, to better understand successful 
project contributions, we must take into consideration and analyze 
different social attributes from the developers creating the commits,
which can be great predictors for the success of a contribution.

Over the years, the open source
community has developed various techniques to vet project contributions, such as pull
requests which must be approved by a core team member~\cite{gousios14}. 
These kinds of mechanisms benefit in substancial part by knowing into which 
aspects to look for as predictors of the quality of the contribution being made.

In this study we analyze a large set of technical and social attributes 
and how these can be used to
add certainty to the probability of a contribution being successful or not. 
Our corpus is based on the entries of a large number of open source projects
using continuous integration. Continuous integration automatically checks
the contribution against the project's test cases, to notify the development
team if the change caused a test case failure~\cite{Meyer14}. The
projects using continuous integration provide quality standard barriers such as
test cases which will cause the build to break if the commit does not perform
accordingly to the standards. When a change
passes these vetting techniques, it shows that the project developers
have confidence in the changes, therefore, allowing us to analyze the success
of the contributions made by passing the quality standards created by the 
core developers of each project.

In past studies, both technical factors and social factors have been found to be
important when predicting pull request success~\cite{tsay14icse,tsay14fse}.
In this study we are interested in analyzing the quality of a contribution 
at a lower granulairity level, commit level. We will analyze this in the 
context of continuous integration, using some of the social attributes provided
by Github as a proxy for the developer's experience level and familiarity with
the process, and their importance when grouped with technical attributes of the commit as 
confounds for the complexity of the commit being pushed.

@@

\textbf{Related work:}
Similar studies have shown the relationship between different attributes and 
different measures of success: pull request acceptance~\cite{gousios14, 
gousios15,tsay14icse,tsay14fse}, pull request evaluation latency~\cite{Yu15}.
Tsay et al~\cite{tsay14icse} performed a very relevant study where they 
determine that the social connection between the developer who is asking the 
pull request and the integrators is a key factor torwards acceptance of a pull 
request. Unlike this previous study, in our study we look at a broader aspect of 
social aspects and how this attributes towards the success in Travis
CI jobs.\footnote{https://docs.travis-ci.com/user/customizing-the-build} 
Gousios et al~\cite{gousios14} study the relationship between the area of the 
source code being modified by the developer requesting the changes to be pulled 
and the acceptance rate of these changes is highly correlated. Also existence of 
tests and overall code quality~\cite{gousios15} are attributes that have a 
strong relationship with the acceptance rate of a pull request. Bettenburgh et 
al~\cite{bettenburgh10} discuss the impact of social interactions with 
post-release defects, Vasilescu et al~\cite{vasilescu14} explored how different 
project characteristics (programming language, project age, etc.) affect commit 
success when integrated to the automatic build process. Yu et al~\cite{yu16} 
analyze the impact of CI usage in Github and its correlation with software 
quality. Unlike the works previously listed, our study focuses on independent social 
attributes and how these relate to the success of the integration of a commit 
pushed by the developer being described by these social attributes.

Our study shows that social factors are an important aspect of build success but
these factors alone do not fully predict if a contribution will be successfully integrated into the project. We also show
that core team members are more likely to achieve build success than non-core
team members and having followers increases the likelihood passing the
continuous integration build.

@@

\subsection{Goal}
The goal of this study is to have a deeper understanding of the relationship 
between the several different social properties that estimate the
developer's experience, and the likelihood of a developer's commit
to be considered successfully integrated in the project. We define successful
integration as passing the continuous integration build tests in the project.

@@

Previous studies have identified a relationship between technical and social 
factors and pull request acceptance. We are interested in understanding if
social factors, used as a proxy for experience, can be used to predict
a different success measure: commit integration success, therefore giving
developers and integrators another tool to be able to identify the quality
of a contribution performed by a particular developer.

@@

\subsection{Metric}
We created a decision tree based on two different sets of data. Merely 
technical data related to technical aspects of the complexity of the commit, and 
a superset of this before-mentioned data and social data gathered from the 
developer who pushed the commit. We evaluate the predictive power of these two 
models with 10 fold cross validation and we analyze which of them is able to 
predict with higher accuracy the success of unseen commits.

\section{Building Dataset}
\label{approach}
To investigate our research questions, we built a data set based on the 
attributes that consisted of both technical and social factors. For the technical factors, we collected
these from the data set provided by the MSR Challenge~\cite{msr17challenge}.
We extracted only the attributes in the data set that proxy the commit's
complexity and the build status as listed below:
\begin{itemize}
%\item tr\_original\_commit
%\item gh\_build\_started\_at 
\item git\_diff\_src\_churn 
\item gh\_diff\_files\_added 
\item gh\_diff\_files\_deleted 
\item gh\_diff\_files\_modified 
\item gh\_diff\_tests\_added 
\item gh\_diff\_tests\_deleted 
\item gh\_diff\_src\_files 
\item gh\_sloc 
\item tr\_status 
\end{itemize}

\begin{table*}[ht]
\centering
\begin{tabular}{l||r|r||r|r}
  & \multicolumn{2}{|c||}{Original number of instances}  & \multicolumn{2}{|c}{Downsampled} \\
\hline
  & Technical & Technical and Social & Technical & Technical and Social \\
\hline
\textbf{Correctly Classified Instances}&  77.01\% & 81.10\%  & 64.10\% & 76.09\%\\

@@

\hline
\end{tabular}
\\
\center
  \caption{ Second and third columns: Comparison of performance when running 10 fold cross validation of two decision trees. The first column after the first double bar uses only technical attributes to predict the Travis Job status, the second column augments the data set with social attributes. Fourth and fifth columns: The same intuition as the previous two, with the difference that the data set was downsampled to provide the same number of successful and unsuccessful instances }
  \label{resultsTable}
\end{table*}

As a result of extracting just these particular technical attributes, the new data set is conformed by a  big number of repeated rows, which would slow down the process of collecting the social data for each of these rows, therefore, at this point we removed the repeated rows and continued to collect the social attributes from the developers who pushed the commits.

To collect the social data, we created a web scrapper to extract the social data
from developer's profiles.
We used the commit hash and the project name in the original data
set provided by the MSR Challenge~\cite{msr17challenge} to locate the web page of the commit using the syntax
below. 
\newline \newline
\textit{https://github.com/\textbf{NameOfProject}/commit/\textbf{CommitHash}}
\newline \newline

Then, we scraped the commit author's
\textit{username} from the page. If the scrapper was able to find username, the
scraper went to the user's profile information using the syntax below.
\newline \newline
\textit{https://github.com/\textbf{UserName}}
\newline \newline
The scrapper then extracted the following items from the user's profile.
\begin{itemize}
\item Repositories
\item Stars
\item Followers
\item Following
\item Contributions in the last year
\end{itemize}

The data set used in this study~\cite{msr17challenge} is very diverse 
and consists of commits that date from far back, and the social data we are 
gathering is recent data, gathered in January 2017. The researchers of this 
project acknowledge this discrepancy and in an effort to mitigate the threat of
using future result to predict previous events,
we removed the commits in the data set that were not performed within the last
three years of the data set's release (2014,2015,2016).  While this is not
perfect, we believe that these numbers are unlikely to have changed
significantly in the last three years.

After building the data set, we converted the various Travis build statuses to
successful (passed in the data set) or not successful (errored, failed, canceled
and started in the data set).  This still left a big discrepancy in the two
categories, with about a 3:1 successful to not successful ratio. To avoid having
this discrepancy alter our results, we created a separate data set where the
successful builds were randomly selected until the number of successful builds
equaled the number of not successful builds (downsampling process).  We used both the
downsampled and non-downsampled data sets in our results.

We evaluated the predictive power of the social factors using a decision
tree~\cite{Quinlan86}.
Decision trees are a predictive model which uses binary decisions on the
attributes to predict a final class.  The predicted class is represented as the
leaves, while the attribute decisions used in the prediction cause the
different branches.
In this study, we used the Weka J48 decision tree~\cite{Weka,Quinlan1993}
with default settings. The decision trees were evaluated using ten fold cross 
validation with the data in each data set.

\subsection{Success Classification}

@@

Table~\ref{resultsTable} shows the results of running the two data sets through the
default Weka J48 decision tree algorithm. The decision tree built from
only the technical indicators has a correct classification rate of 77.01\% which
means the technical attributes only increased the prediction accuracy by about 3\%.
Adding the social data increased the classifier's accuracy from 77.01\% to
81.10\%, about a 4\% increase.

@@

\section{Discussion}
We have found that social attributes which proxy experience produce a 
noticeable increase in predicting build success.  While the increase was noticeable 
with the full data set, there was a much greater increase in prediction
accuracy after performing downsampling. These results show that social factors
as the ones studied in this paper are important predictors when
evaluating the predictive power of build commit success.

The build success by core team member results show that core team members are
more likely to have the build integration pass than non-core team members.  This
result provides evidence that non-core team member commit are not thoroughly
scrutinized, although further studied will be required to draw definite
conclusions.

@@

\section{Threats to validity}
\noindent\textbf{Internal validity:}\\
The researchers in this study used the dataset provided by the 
 MSRChallenge~\cite{msr17challenge} to gain useful information regarding the 
 behavior of developers when pushing commits and how well these commits integrate 
 successfully with the project. The authors have used 
 Weka,\footnote{http://www.cs.waikato.ac.nz/ml/weka/} the well known tool for 
 usage of machine learning algorithms to analyze our datasets. We are aware that, 
 as every human written tool, Weka is prone to contain errors and/or bugs; 
 nonetheless, Weka is a broadly used, open source tool that is subject to high 
 quality standards and a vast user and developer base which makes it more likely 
 to be a high quality code and to have less bugs than lesser used tools.

 The scripts we have created to adapt and extract the datasets we needed, and the 
 scripts to obtain the developer social data are publicly available to facilitate 
 the transparency and quality of the process(Will be made available after double blind review).%\footnote{https://goo.gl/CqTX44}

\noindent\textbf{External validity:} \\
 To tackle the concern that the results of this study might not generalize to 
 other instances outside of the studied cases, we have used a broad and rich 
 dataset~\cite{msr17challenge} in our analysis that comprises 1,359 projects. The 
 projects used in this dataset have also been filtered to be popular (having more 
 than 10 watches on Github), and to have a history of using the continuous 
 integration tool Travis CI for more than 50 builds. All these features help to 
 build the notion that since we are building our analysis on top of a vast 
 dataset comprehending a wide diversity of projects, the threat of this analysis 
 not generalizing to other projects is mitigated.

\noindent\textbf{Construct validity:}\\
The authors understand that 
 Travis job success is not the only way to analyze the success of a commit, and 
 that for different reasons the job may not be successful even if the code 
 commited does not have flaws (a flacky test case for example). Since continuous 
 integration is a best known practice in industry and the use of this methodology 
 spreads out into the open source community, we consider that cases of misuse of 
 CI such that it will lead to failing travis jobs when they should have actually 
 be successful are minimum, specially since having a solid build is one of the 
 main purposes of the methodology, and taking into account that the projects in 
 out dataset are projects with a history of using continuous integration and 
 having a certain level of popularity, this threat is minimized. We therefore 
 consider Travis job success to be a proxy 
 for success when integrating a commit into a larger project.

\section{Conclusion}
In this study we have analyzed the probability of Travis job success as being predicted by two different categories of attributes: Technical and Social data.
We gathered the attributes that describe complexity of each commit being pushed and have created a decision tree that predicts if a future pushed commit will be successfully integrated into its corresponding build or not.
Afterwards, we have gathered social attributes from the developers who pushed each of these commits
and we have augmented our dataset with these attributes to understand whether adding social information about the developer would help the decision tree predict with higher accuracy future 
instances.

Our results show that adding social data to the construction of the decision tree increases the probability this correctly predicting future commit's success in integrating into their corresponding projects by a margin of 4.09\% to 11.99\%, depending on the ratio of successful vs unsuccessful instances are used to train the desicion tree.

We have also found that non-core contributors cause not successful builds more
than core contributors.  This leads us to believe that non-core contributions
are not inspected as closely as they should be.  We have also found that having
a follower, and especially a large follower base, can be an indicator of build
integration success.
